---
title: "Psych 215 - SP26"
format: html
editor: visual
navbar: false
---

## Research Methods in Psychology

```{r}
#| echo: false
#| message: false
#| warning: false

library(googlesheets4)
library(gt)
library(tidyverse)
options(knitr.kable.NA = "")
gs4_deauth()
d <- read_sheet("https://docs.google.com/spreadsheets/d/1OegmH097biy7yzrxX98NgkXa04WU65EF4e4xwm0rusI/edit?usp=sharing", sheet = "215")
d$Date <- format(d$Date, "%m/%d")
```

Hi class! This page will contain deadlines and links to course readings and lecture notes.

**Please use Canvas to :**

-   submit Reading Quiz, Discussion Posts, and Project Assignments.

-   access the Syllabus, Lecture Recordings, and our class Vision Board.

## Course Materials

-   [Dig Deeper Description & Rubric](/rm/notes/digdeeper.html)

-   Final Project Description & Rubric

## Course Outline

Note : Chapters are from the open-source textbook : [Research Methods in Psychology (4th Edition)](https://kpu.pressbooks.pub/psychmethods4e/). Note that each chapter has several sections within it that you'll need to click through. Here's a link to [download a PDF](https://kpu.pressbooks.pub/psychmethods4e/open/download?type=pdf) of the textbook.

```{r}
#| echo: false
knitr::kable(d, padding = 5)
```
