---
title: "Linear Models Practice Script"
format: 
  html:
      code-overflow: wrap
---

## The Linear Model in FOUR EASY STEPS.

The model is a line that updates our predictions of one variable based on knowledge of another.

1.  Load your data and check to make sure your data loaded correctly.

```{r}
#| include: false
m <- read.csv("https://www.dropbox.com/scl/fi/wczjc24dxptfcgdgdijaj/mini_DATA.csv?rlkey=83vd2ulnxgbd8brmip0f41txq&dl=1", stringsAsFactors = T)
head(m)
nrow(m)
```


2.  Graph your DV and IV(s) : make sure the data look good (remove any outliers) and think about these variables - how do you think they will be related?

```{r}

```

3.  Plot the relationship between the two variables, define the linear model, and add this line to your plot. 

```{r}

```

4.  Then, interpret the intercept, slope, and error ($R^2$) of the model. What do these terms tell you about the relationship between the two variables?

```{r}

```

5.  Finally, use your human brain to think about this relationship. Why does it exist? What do you learn? What do you not know? What other questions do you have?

