---
title: ""
format: html
---

# Faking Data Is Okay As Long As You Are Honest

## ACTIVITY : VISUALIZING SAMPLING ERROR 

### STEP 1. Be an omnipotent power who can create an entire world of individuals.

```{r}
# rnorm(10000000, mean = 100, sd = 20)
fakey <- rnorm(10000000, mean = 100, sd = 30)
length(fakey)
head(fakey)
```

### STEP 2. Run some stats on these data as we do.

```{r}
mean(fakey, na.rm = T)
hist(fakey)
abline(v = mean(fakey), lwd = 5)
```

### STEP 3. Take a random sample from this population.

The sample function can be used to take a "perfectly random" sample from a set of data.

```{r}
?sample # our friend, the sample function
sample(1:10, 1) # are you vibing with R?
sample(1:10000000, 1) # are you REALLY vibing?
sample(1:length(fakey), 1) # the numbers to sample from, another way. why better?
sample(1:length(fakey), 10) # a small sample

```

We can adapt this sample function to draw a random sample from fakey, using indexing.

```{r}
fakey[3] # indexes person 3 from fakey
fakey[sample(1:length(fakey), 10)] # ten random individuals from fakey.
```

Every time we run this code, we will generate a "new" random sample.
```{r}
fakey[sample(1:length(fakey), 10)]
```

So if we want to "save" a random sample, we need to save it as an object.
```{r}
lilfakey <- fakey[sample(1:length(fakey), 10)] # ten random individuals from fakey.
lilfakey
```

### STEP 4. Run some stats on this sample.

```{r}
mean(lilfakey)
hist(lilfakey)
abline(v = mean(lilfakey), lwd = 5)
```

### STEP 5. Repeat These Steps Over and Over Again

We could run the lines below over and over and over again to illustrate sampling error...

```{r}
lilfakey <- fakey[sample(1:length(fakey), 10)] # ten random individuals from fakey.
hist(lilfakey, xlim = c(0,200), ylim = c(0,10),
     breaks = 5,
     main = paste(c("mean=", round(mean(lilfakey), 4)), sep = ""))
abline(v = mean(lilfakey), lwd = 5)
```

"There's got to be a better way."

### STEP 5 : THE FOR-LOOP.

We can ask R kindly to do this repetition 1000 times for us.

```{r}
truthbucket <- array()
for(i in c(1:1000)){
  lilfakey <- fakey[sample(1:length(fakey), 10)] # ten random individuals from fakey.
  truthbucket[i] <- mean(lilfakey)
}
length(truthbucket)
hist(truthbucket)
mean(truthbucket)
```

### QUESTION : WHAT STATISTIC(S) COULD DESCRIBE HOW MUCH ERROR THERE IS IN OUR SAMPLING???????

```{r}

```

### The 95% Confidence Interval.
```{r}
hist(truthbucket)
abline(v = mean(truthbucket), lwd = 5)
abline(v = mean(truthbucket) + 1.96 * sd(truthbucket), lwd = 2, col = 'red')
abline(v = mean(truthbucket) - 1.96 * sd(truthbucket), lwd = 2, col = 'red')
```

### ACTIVITY : How does the sampling error change when our sample size increases?

```{r}

```

# Bootstrapping : Our Data

We'll start with a sample of data

```{r}
d <- read.csv("https://www.dropbox.com/scl/fi/l2vhaw1p4kkng7vozgv8f/DATASET_happy_data.csv?rlkey=wg1buadbuuv712ia55ju7lk0s&dl=1", stringsAsFactors = T)
head(d)
hist(d$Child.Mortality)
mean(d$Child.Mortality, na.rm = T)

```


And we can use a for-loop to estimate sampling error

```{r}
sample(1:nrow(d), nrow(d), replace = T)
d2 <- d[sample(1:nrow(d), nrow(d), replace = T),]
hist(d2$Child.Mortality, xlim = c(0,15), ylim = c(0,70))
abline(v = mean(d2$Child.Mortality, na.rm = T), lwd = 5)
mean(d2$Child.Mortality, na.rm = T)
```
But the for-loop does this many times.

```{r}
worldbucket <- array()
for(i in c(1:1000)){
  d2 <- d[sample(1:nrow(d), nrow(d), replace = T),]
  worldbucket[i] <- mean(d2$Child.Mortality, na.rm = T)
}
```

And we can look at the results of this resampled distribution.

```{r}
hist(worldbucket, xlim = c(0,5))
mean(d$Child.Mortality, na.rm = T)
mean(worldbucket)
sd(worldbucket) # SAMPLING ERROR.
```

## evaluate our estimate, relative to the sampling error

```{r}
upper <- mean(d$Child.Mortality, na.rm = T) + 1.96*sd(worldbucket) # SAMPLING ERROR.
lower <- mean(d$Child.Mortality, na.rm = T) - 1.96*sd(worldbucket) # SAMPLING ERROR.
hist(worldbucket, xlim = c(0,5))
abline(v = upper)
abline(v = lower)
```
