---
title: "Lecture 8 | Interaction Effects"
format: 
  html:
    code-overflow: wrap
---

## Announcements

-   Lab 8 : Self-Care Lab. Do something nice for yourself you wouldn't have done if you had a lab. Share on discord (or not!)
-   Final Project : A Discussion Post; identify a dataset and research question & think about how you might answer this question with the data (and a linear model??)
-   Our Remaining Class Time
    -   4/4 Class : On Zoom or Wizard of Oz Style?? (And section??). Topic : Logistic Regression and Quadratic Terms and Other Linear Model Stuff.
    -   4/11 : IRL, on Hierarchical Linear Models.
    -   4/18 : MEGA Exam. Will do my best to keep it chill.
    -   4/25 : More Hierarchical Linear Models. PCA? We will see!!
    -   5/2 : Our last class can u believe it???? The learning has stopped. So review, discussion, tears, project time.

## [Check-In : Interpreting a Multiple Regression Table.](https://docs.google.com/forms/d/e/1FAIpQLSdEZdIHsPohB1U3iVcp_PrlGPjXPDr-UyozAlHAbw6n2YlS4w/viewform?usp=header)

Note : these data come from the "hormone_datset.csv"; in the dropbox folder. But not necessary to download?

```{r}
#| include: false
#install.packages("jtools")
library(jtools)
library(arm)
#h <- readxl::read_xls("~/Downloads/in.excel.for.adcat.hormone.data.xls")
#h <- h[,-1]
#write.csv(h, "~/Dropbox/!GRADSTATS/gradlab/Datasets/hormone_dataset.csv", row.names = F)
h <- read.csv("~/Dropbox/!GRADSTATS/Datasets/hormone_dataset.csv")
h$sexF <- as.factor(h$sex)
levels(h$sexF) <- c("Male", "Female")
h$sexF <- relevel(h$sexF, ref = "Female")

hmod <- h[c(!is.na(h$test_mean)), ]



#zm1 <- scale_mod(mod1, scale.response = T)
#zm2 <- scale_mod(mod2, scale.response = T)
#zm3 <- scale_mod(mod3, scale.response = T)
```

```{r}
mod1 <- lm(narcicissm ~ sexF, data = hmod)
mod2 <- lm(narcicissm ~ test_mean, data = hmod)
mod3 <- lm(narcicissm ~ sexF + test_mean, data = hmod)
export_summs(mod1, mod2, mod3, 
             coefs = c("Sex (0 = Female; 1 = Male)" = "sexFMale", "Testosterone" = "test_mean"),
             error_format = "[se = {std.error}, t = {statistic}]", transform.response = T, scale = T, confint = T)
```

## Multiple Regression : More Regression

### Labeling Changes

1.  "Independent" Effect. The slope of the IV does not change when other variables are added to the model.
2.  "Suppression" Effect. The slope of the IV gets stronger (in either direction) when other variables are added to the model.
3.  "Mediation" Effect. The slope of the IV gets closer to zero when other variables are added to the model. *Careful : "mediation" often implies some causal relationship, and it is very hard (/impossible?) to estimate causal relationships without experimental methods. Some links below that go deeper into this.*

### How Large a Change is Enough to Matter???

Is the difference in slope between Model 1 and Model 3 significant?

```{r}
export_summs(mod1, mod2, mod3, 
             coefs = c("Sex (0 = Female; 1 = Male)" = "sexFMale", "Testosterone" = "test_mean"),
             error_format = "[se = {std.error}, t = {statistic}]", transform.response = T, scale = T, confint = T)
```

```{r}
bucket <- array()
for(i in c(1:1000)){
  hsim <- h[sample(1:nrow(h), nrow(h), replace = T),]
  xm1 <- lm(narcicissm ~ sexF, data = hsim)
  xm2 <- lm(narcicissm ~ sexF + test_mean, data = hsim)
  bucket[i] <- coef(xm1)[2] - coef(xm2)[2]
}
sum(bucket > 0)
```

```{r}
export_summs(mod1, mod2, mod3,
             coefs = c("Sex (0 = Female; 1 = Male)" = "sexFMale", "Testosterone" = "test_mean"),
             error_format = "[se = {std.error}, t = {statistic}]", 
             transform.response = T, scale = T, confint = T)
```

```{r}
# install.packages("mediation")
library(mediation)
modM <- lm(test_mean ~ sexF, data = h) # defining a model to predict our mediator (testosterone) from the IV (sex)
medmod <- mediate(modM, mod3, treat = "sexF", mediator = "test_mean", dropobs = "TRUE", boot = TRUE, sims = 1000)
summary(medmod)
```

-   Total Effect : slope of sexFMale in Model 1
-   Average Direct Effect : slope of sexFMale in Model 3
-   ACME = Average Causal Mediation Effects = Total Effect - Direct Effect; influence our mediatior has on the pre-existing relationship.

**NOTE : the output here doesn't match the effects in the table for two reasons:**

1.  My results in the table are standardized (and `mediate` doesn't seem to work well with z-scored models? I couldn't figure it out at least...)
2.  `mediate` will only include complete cases across all the models. Notice that my first linear model has a larger sample size than the models with testosterone. It would be a more fair comparison to remove the individuals from Model 1 who are also missing in my other models. Will chat more about missing data next lecture.

### Would You Like to Learn More?

-   [The Difference Between “Significant” and “Not Significant” is not Itself Statistically Significant](https://sites.stat.columbia.edu/gelman/research/published/signif4.pdf)
-   [A conceptual review of mediation, and tour through the 'mediation' package.](https://library.virginia.edu/data/articles/introduction-to-mediation-analysis)
-   Preacher KJ, Hayes AF. Asymptotic and resampling strategies for assessing and comparing indirect effects in multiple mediator models. Behav Res Methods. 2008;40(3):879-891. doi:10.3758/BRM.40.3.879
-   [A nice and recent summary of why caution is needed for mediation analysis.](https://www.the100.ci/2025/03/20/reviewer-notes-thats-a-very-nice-mediation-analysis-you-have-there-it-would-be-a-shame-if-something-happened-to-it/)

## [Presentations](https://docs.google.com/presentation/d/1YZQ45_oj6TgiSIUpU7N6Ek5iTk2nn4T5RIpCz6Xi1K4/edit?usp=sharing)

## Multiple Regression : Moderation / Interaction Effects

### Frantz Fanon Reading.

1.  What did you think about the reading???
2.  What is his main point?
3.  Have you observed this in the wild / our modern society?
4.  How would or might you test this idea with data, as a modern psychologist would???
5.  Any other questions / ideas / comments?

### [See Professor Handout.](https://docs.google.com/document/d/1YVzL5ud5UGy82KXqGrc-mjSRFJh7ocuQpXdIBmbvd0o/edit?usp=sharing)

### Picture is worth...that's right class...1000 words.

```{r}
library(ggplot2)
library(ggthemes)
ggplot(data = h, aes(x = scale(test_mean), y = scale(narcicissm), color = sexF)) + 
  geom_point(size = .5, alpha = .3, position = "jitter") + 
  labs(title = "The Interaction Effect", x = "Testosterone (Z-Scored)", y = "Narcissism (Z-Scored)", color = "Sex") + 
  geom_smooth(method = lm) + theme_apa()
```

### Still just a linear model...

A linear model whose IVs *must* be standardized!!

```{r}
mod4 <- lm(narcicissm ~ sexF * test_mean, data = h)
m4z <- scale_lm(mod4, scale.response = T) # you 
m4z
```

```{r}
cf4 <- coef(m4z)
plot(scale(narcicissm) ~ scale(test_mean), col = sexF, data = h, pch = 19)
abline(a = cf4[1],
       b = cf4[3], 
       col = "black", lwd = 5) # line for females
abline(a = cf4[1] + cf4[2],
       b = cf4[3] + cf4[4], 
       col = "red", lwd = 5) # line for males
```

### Reporting, and in a table.

```{r}
ggplot(data = h, aes(x = scale(test_mean), y = scale(narcicissm), color = sexF)) + 
  geom_point(size = .5, alpha = .3, position = "jitter") + 
  labs(title = "The Interaction Effect", x = "Testosterone (Z-Scored)", y = "Narcissism (Z-Scored)", color = "Sex") + 
  geom_smooth(method = lm) + theme_apa()
export_summs(mod1, mod2, mod3, mod4,
             coefs = c("Sex (0 = Female; 1 = Male)" = "sexFMale", "Testosterone" = "test_mean", "Sex * Testosterone" = "sexFMale:test_mean"),
             error_format = "[se = {std.error}, t = {statistic}]", transform.response = T, scale = T, confint = T)
```

### Would You Like To Learn More?

As always, let me know if you find great resources :)

-   [The Interactions Package](https://interactions.jacob-long.com). The author of jtools has made what looks like a super clean way to graph interaction effects in R, using ggplot2.
-   [Visualization Tutorials](https://sebastiansauer.github.io/vis_interaction_effects/). Other, more cumbersome methods of plotting interaction effects exist!
