---
title: "Research Methods : Statisitcs Review"
format: html
---

## Loading (and Learning About) Our Dataset

These data come from a survey of our class; a study of 40ish community college students who are taking a research methods class as part of a degree requirement.

The survey measured students' attitudes about reading through self-report questions.

A list of variables is below.

```{r}
#| echo: false

## Data Sources
dataurl <- "https://www.dropbox.com/scl/fi/47gw0h2x6abprskwdqgo7/DATASET_ReadingData.csv?rlkey=4mm3itw9p4fm7ee5qa7z3n82q&dl=1"
codebookurl <- "https://www.dropbox.com/scl/fi/m4os7p2jualfzzuyqbnqy/CODEBOOK_ReadingData.csv?rlkey=wteyc27f8vgo6ihe05hijnkln&dl=1"

## Defining the Dataset and Codebook
d <- read.csv(dataurl, stringsAsFactors = T)
codebook <- read.csv(codebookurl)

## Visualizing the Codebook
knitr::kable(codebook)
```

## Graphing Variables.

The hist() function graphs a numeric variable.

The code below asks R to define a histogram, using the dataset d, of the variable like.

```{r}
hist(d$like)
```

::: {.callout-tip collapse="true"}
### Arguing with R

We can change some of the *arguments* of the hist() function by adding these instructions to the code above to make the graph look nicer.

-   `main = "new title go here" # changes the title of the graph`

-   `xlab = "new name go here" # changes the name of the x-axis label`

-   `ylab = "new name go here" # changes the name of the y-axis label`

-   `col = "blue" # changes the color of the graph to blue (or any color)`

-   `xlim = c(1,5) # changes the range of the x-axis to go from 1 to 5 (or any number you want)`

-   `ylim = c(0,20) # changes the range of the y-axis to go from 1 to 20 (or any number you want)`
:::

### Describing Variables with Human Brains

What do you see about the graph above? No stats language allowed.

#### What the Graph is Showing

-   student

-   ideas

-   go

-   here

#### Things I Learn About the Graph

-   student

-   ideas

-   go

-   here

#### Questions That I Have / Things I Do Not Know

-   student

-   ideas

-   go

-   here

### Statistics as a Language

#### Defining Variables With Numbers

| Statistic | What it Describes | R Code | Result for `d$like` | What we Learn From This Statistic |
|---------------|---------------|---------------|---------------|---------------|
| mean | the value that is closest to all the other scores in the dataset; a prediction about the expected value | `mean(d$V1)` | `r round(mean(d$like), 2)` |  |
| median | the value that is in the middle of all the scores (50% of scores are above the median, 50% of scores are below the median) | `median(d$V1)` | `r round(median(d$like), 2)` |  |
| range | the minimum and maximum value of the data | `range(d$V1)` | `r round(range(d$like), 2)` |  |
| standard deviation | how much the average individual differs from the mean (or : the average "error" when we use the mean to make a prediction) | `sd(d$V1)` | `r round(sd(d$like), 2)` |  |

: {tbl-colwidths="\[15,30,15,10,30\]"}

### Discussion : How Can We Connect our "Brain Learnings" To These Numbers?

### ACTIVITY : The Learning Will Continue

1.  Choose a variable from the READING dataset that you are intersted to learn more about.
2.  Type R commands in the "chunk" below to graph this variable. Try to change the arguments to make the graph look nice.
3.  Below the graph, type R commands to calculate descriptive statistics for this variable. Connect these to the graph.
4.  What do you learn about this variable from the graph?
5.  Share your graph, the statistics, and your learnings in our class vision board.

```{r}

```
